---
title: "Climate Modelling"
author: "Pietro Monticone & Davide Orsenigo"
date: '`r Sys.Date()` | Turin University'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
devtools::install_github("yihui/xaringan")
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)

csss508css <- list(
     ".hljs-tomorrow-night-bright .hljs" = list(
        "background" = "#10102c",
        "border-radius"="5px"),
     ".remark-inline-code" = list(
        "background" = "#E0E0E0",
        "color" = "#10102c",
        "border-radius" = "3px",
        "padding" = "2px"),
     ".inverse .remark-inline-code" = list(
        "background" = "#10102c",
        "color" = "#ececf8",
        "border-radius" = "3px",
        "padding" = "2px"),
     ".small" = list("font-size" = "75%"),
     ".smaller" = list("font-size" = "60%"),
     ".remark-code-line-highlighted" = list(
        "background-color" = "rgba(255,0,255,0.2)"),
     "sup" = list("font-size" = "14px"),
     "h1, h2, h3" = list(
        "margin-top"=".25em", 
        "margin-bottom"=".25em"),
     ".pull-left60" = list(
        "float" = "left",
        "width" = "58%" ),
     ".pull-right40" = list(
        "float" = "right",
        "width" = "38%" ),
     ".pull-right30" = list(
        "float" = "right",
        "width" = "30%" ),
     "a" = list("text-decoration" = "underline"),
     ".inverse a" = list("color" = "#cbd3a3"),
     "body" = list("line-height" = "1.4")
    )

mono_accent(base_color = "#342c5c",
            code_font_google = google_font("Fira Mono"),
            header_font_google = google_font("Quattrocento"),
            extra_css = csss508css,
            title_slide_background_position = "bottom",
            title_slide_background_size = "contain"
)
```
# Contents 
- Introduction
  - Packages
  - Input Parameters
  - Fit from Textbook Spreadsheet
  - Functions
- Run 0
- Run 1
- Run 2
- Run 3 & 4
- Hysteresis Cycles
- Embedding Daisies in EBM

# Packages
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(tidyverse)
library(plotly)
library(scales)
```
---
# Input parameters

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
S <- 1370 
A <- 204
B <- 2.17
K <- 3.86
ai <- 0.62
ab <- 0.25
Temperature <- rep(0,300)
Temp <- rep(0,100)
Temp2 <- rep(0,180)
Temp5 <- rep(0,300)
J <- rep(0,100)
Albedo <- rep(0,300)
t <- c(1:300)
Zones <- seq(-89, 89, by = 2)
```
---
## Fit from Textbook Spreadsheet
.small[
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE,out.width="50%"}
zones <- c(85,75,65,55,45,35,25,15,6)
coszones <- cospi(zones/180)
sunWt <- c(0.5,0.531,0.624,0.77,0.892,1.021,1.12,1.189,1.219)
df <- data.frame(zones,sunWt)
f <-  function(zones,d,n,k){ d*cos(n*zones)^2+k }
Fit <-  nls(sunWt ~ f(zones,d,n,k),data= df, start=list(d=1,n=0.015,k=0.3))
ggplot(df, aes(zones,sunWt))+geom_point()+geom_smooth(aes(y=f(zones,0.7768699,0.0164348,0.4617747)))+xlab("Latitude")+ylab("Solar Weight Factor")
```
]
---
# Functions
.pull-left[
 .small[
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
Func <-  function(x){0.7768699*cos(0.0164348*x)^2+0.4617747}
gauss <- function(x,m,sd,b){
          ((24+b)/(0.00798*sqrt(2*pi*sd^2)))*exp(-(x-m)^2/(2*sd^2))-b}
Sun1 <- function(x,a){a*x/100}
Sun2 <- function(x){1370*(sinpi((x+90)/180))^2}
Sun3 <- function(x){
          1370-(((1370)/(sqrt(2*pi*1^2)))*exp(-(x-50)^2/(2*1^2)))}
Sun4 <- function(x){ifelse(x==50,1370/3,1370)}
Sun5 <- function(x){(1/100) * (abs(x-150)+25)}
Sun6 <- function(x){1370*(1+0.1*cospi(x/180))}
Incident <- function(x,y){x*y/4}
Step <- function(x,c){ifelse(x<c, 0.6, 0.3)}
alb <- function(x,a,b){(-exp(2.2*x+10)/(exp(2.2*x+10)+1))*(a-b)+a}
```
 ]
]

$$ S_1(t)= \frac{1370}{100}x $$
$$ S_2(t)= 1370(\sin^2(x+90)) $$
$$ S_3(t)= 1370\Big(1-\frac{1}{\sqrt{2\pi}}e^{-(x-50)^2/2}\Big) $$
$$ S_4(t)=1370\Big(1-\frac{1}{3}\delta(t-50)\Big) $$
$$ S_5(t)=\frac{1}{100}(|x-150|+25) $$
$$S_6(t)=1370\Big(1+\frac{1}{10} cos(x)\Big) $$
---
# Introduction
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
Ti <- gauss(Zones,0,50,31.6)
cosZones <- abs(cospi(Zones/180))
SunWt <- Func(Zones)
Rin <- Incident(S,SunWt)
T <- gauss(Zones,0,50,31.6)
a <- alb(T,ai,ab)
```
--
# Run 0
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
for(i in c(1:300)) {
  Tcos <- cosZones*T
  Tm <- sum(Tcos)/sum(cosZones)
  T <- (Rin*(1-a)+K*Tm-A) / (B+K)
  a <- alb(T,0.62,0.25)
  Temperature[i] <-  Tm 
} 
```
---
.small[
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
data1 <- data.frame(Zones,T,a,Ti)
plot1 <- ggplot(data1)+geom_line(aes(Zones, T),colour = 'red')+geom_line(aes(Zones, a*25), colour = 'blue')+geom_line(aes(Zones, Ti), colour = 'red', linetype = "dashed")+scale_y_continuous(sec.axis = sec_axis(~./25,name = "a"))+xlab("Latitude")+ylab("Temperature")
```
]
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
plot1
```
---
.small[
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
data2 <- data.frame(t,Temperature)
plot2 <- ggplot(data2) +
  geom_line(aes(t, Temperature),colour = 'green')+xlab("Time")+ylab("Mean Temperature")
```
]
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
plot2
```
---
# Run 1
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
TEMP1 <- matrix(NA, nrow=90, ncol=100)
for(j in c(1:100)){
  T <- gauss(Zones,0,50,31.6)
  a <- alb(T,ai,ab)
  S <- Sun1(1370,j)
  Rin <- Incident(S,SunWt)
     for(i in c(1:300))
     {Tcos <- cosZones*T
      Tm <- sum(Tcos)/sum(cosZones)
      T <- (Rin*(1-a)+K*Tm-A) / (B+K)
      a <- alb(T,ai,ab)}
  TEMP1[,j] <- T
  J[j] <- j
  Temp[j] <- Tm }
```
---
.small[
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
data3 <- data.frame(J,Temp)
plot3 <- ggplot(data3,aes(J,Temp))+geom_line(aes(J, Temp),colour = 'red')+xlab("Time")+ylab("Mean Temperature")+ggtitle("Non Dynamic Mean Temperature Temperatura (T(t) independent from T(t-1))")
```
]
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
plot3
```
---
.small[
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
plot_ly(z=~TEMP1)%>% add_surface() %>%   layout(
    title = "With Initialization", scene = list(
      xaxis = list(title = "S/100"),
      yaxis = list(title = "Latitude"),
      zaxis = list(title = "T")  )) 
```
]
---
# Run 2
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
T2 <- gauss(Zones,0,50,31.6)
TEMP2 <- matrix(NA, nrow=90, ncol=180)
a <- alb(T2,ai,ab)
Sarr <- rep(0,180) 
for(j in c(1:180)){
  S <- Sun2(j)
  Rin <- Incident(S,SunWt)
  for(i in c(1:300))
  {Tcos <- cosZones*T2
  Tm <- sum(Tcos)/sum(cosZones)
  T2 <- (Rin*(1-a)+K*Tm-A) / (B+K)
  a <- alb(T2,ai,ab)
  }
  Sarr[j] <- Sun2(j)
  TEMP2[,j] <- T2
  J[j] <- j
  Temp2[j] <- Tm
}
```
---
.small[
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
data3 <- data.frame(J,Temp2)
plot3 <- ggplot(data3,aes(J,Temp2))+geom_line(aes(J, Temp2),colour = 'red')+xlab("Time")+ylab("Mean Temperature")+ggtitle("Sinusoidal Perturbation")
```
]
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
plot3
```
---
.small[
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
plot_ly(z=~TEMP2) %>% add_surface() %>% layout(
    title = "Without Initialization", scene = list(
      xaxis = list(title = "S/100"),
      yaxis = list(title = "Latitude"),
      zaxis = list(title = "T")  ))
```
]
---
# Run 3 & 4
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
TEMP3 <- matrix(NA, nrow=90, ncol=200)
T3 <- gauss(Zones,0,50,31.6)
a <- alb(T3,ai,ab)

for(j in c(1:200)){
  S <- Sun3(j)
  Rin <- Incident(S,SunWt)
  for(i in c(1:300))
  {Tcos <- cosZones*T3
  Tm <- sum(Tcos)/sum(cosZones)
  T3 <- (Rin*(1-a)+K*Tm-A) / (B+K)
  a <- alb(T3,ai,ab)  }
  TEMP3[,j] <- T3
  J[j] <- j
  Temp[j] <- Tm
}
```
---
.small[
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
data3 <- data.frame(J,Temp)
plot3 <- ggplot(data3,aes(J,Temp))+geom_line(aes(J, Temp),colour = 'red')+xlab("Time")+ylab("Mean Temperature")+ggtitle("Gaussian Perturbation")
```
]
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
plot3
```
---
.small[
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
plot_ly(z=~TEMP3) %>% add_surface()%>% layout(
    title = "Without Initialization",  scene = list(
      xaxis = list(title = "S/100"),
      yaxis = list(title = "Latitude"),
      zaxis = list(title = "T")   ))
```
]
---
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
TEMP4 <- matrix(NA, nrow=90, ncol=200)
T4 <- gauss(Zones,0,50,31.6)
a <- alb(T4,ai,ab)

for(j in c(1:200)){
  S <- Sun4(j)
  Rin <- Incident(S,SunWt)
  for(i in c(1:300))
  {Tcos <- cosZones*T4
  Tm <- sum(Tcos)/sum(cosZones)
  T4 <- (Rin*(1-a)+K*Tm-A) / (B+K)
  a <- alb(T4,ai,ab)
  }
  TEMP4[,j] <- T4
  J[j] <- j
  Temp[j] <- Tm
}
```
---
.small[
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
data3 <- data.frame(J,Temp)
plot3 <- ggplot(data3,aes(J,Temp))+geom_line(aes(J, Temp),colour = 'red')+xlab("Time")+ylab("Mean Temperature")+ggtitle("Delta Perturbation")
```
]
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
plot3
```
---
.small[
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
plot_ly(z=~TEMP4) %>% add_surface()%>% layout(
    title = "Without Initialization", scene = list(
      xaxis = list(title = "S/100"),
      yaxis = list(title = "Latitude"),
      zaxis = list(title = "T")  ))
```
]
---
# Hysteresis Cycles
---
.small[
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
data <- data.frame(Sarr,Temp2)
plot <- ggplot(data,aes(Sarr,Temp2))+geom_point(aes(Sarr, Temp2),colour = 'yellow')+ylab("Mean Temperature")+xlab("S_2(t)")+ggtitle("Mean temperature vs. Sinusoidally Fluctuating Incoming Radiation")
```
]
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
plot
```
---
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
Temp5 <- rep(0,250)
T5 <- gauss(Zones,0,50,31.6)
a <- alb(T5,ai,ab)
Sarr <- rep(0,250) 
for(j in c(0:250)){
  S <- Sun5(j)
  Rin <- 1370*Incident(S,SunWt)
  for(i in c(1:60))
  {Tcos <- cosZones*T5
  Tm <- sum(Tcos)/sum(cosZones)
  T5 <- (Rin*(1-a)+K*Tm-A) / (B+K)
  a <- alb(T5,ai,ab)
  }
  Sarr[j] <- Sun5(j)
  J[j] <- j
  Temp5[j] <- Tm
}
```
---
.small[
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
dataw <- data.frame(Sarr,Temp5)
plot5 <- ggplot(dataw,aes(Sarr,Temp5,group=1))+geom_point(aes(Sarr, Temp5),colour = 'yellow')+geom_segment(aes(x = Sarr[89], y = Temp5[89], xend = Sarr[90], yend = Temp5[90]),colour = 'yellow')+geom_segment(aes(x = Sarr[258], y = Temp5[258], xend = Sarr[259], yend = Temp5[259]),colour = 'yellow')+ylab("Mean Temperature")+xlab("S_5(t)")+ggtitle("Mean temperature vs. Linearly Fluctuating Incoming Radiation")
```
]
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
plot5
```

---
# Embedding Daisies in EBM
---
## Input Parameters
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
S <- 1370
A <- 204
B <- 2.17
K <- 3.86
ai <- 0.62
ab <- 0.25
aW <- 0.75
aB <- 0.25
c <- 7
k <- 0.003265*0.75
T0 <- 20
D <- 0.3 # Death Rate
Zones <- seq(-89, 89, by = 2)
```
## Functions
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
Func <-  function(x){
  0.7768699*cos(0.0164348*x)^2+0.4617747
}
gauss <-  function(x,m,sd,b){
  ((24+b)/(0.00798*sqrt(2*pi*sd^2)))*exp(-(x-m)^2/(2*sd^2))-b
}
Incident <- function(x,y){x*y/4}
alb <- function(x,a,b){ (-exp(1.6*x+10)/(exp(1.6*x+10)+1))*(a-b)+a} #naked
```
---
## Initialization
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
cosZones <- abs(cospi(Zones/180))
SunWt <- Func(Zones)
Rin <- Incident(S,SunWt)
T <- gauss(Zones,0,50,31.6)-6

w <- rep(0.5,length(Zones))
b <- rep(0.2,length(Zones))
u <- rep(0.3,length(Zones))
a <- w*aW+b*aB+u*alb(T,0.62,0.25)
Barr <- rep(0,500)
Warr <- rep(0,500)
Uarr <- rep(0,500)
Tarr <- rep(0,500)
I <- rep(0,500)
TEMP <- matrix(NA, nrow=90, ncol=500)
```
---
## EBM without Daisies
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
for(i in c(1:500)) {
  S <- Sun6(i)  # oppure costante S <- 1370
  Rin <- Incident(S,SunWt)
  Tcos <- cosZones*T
  Tm <- sum(Tcos)/sum(cosZones)
  T <- (Rin*(1-a)+K*Tm-A) / (B+K)
  TEMP[,i] <- T
  a <- alb(T,0.62,0.25)
  I[i] <- i
  Tarr[i] <- T[45]
} 
```
---
.small[
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
plot <- ggplot(data.frame(Zones,w,b,u,T))+geom_line(aes(Zones,w),color="white")+geom_line(aes(Zones,b),color="black")+geom_line(aes(Zones,u),color="brown")+geom_line(aes(Zones,T/5),color="green")+ylab("Temperature & Albedo")+xlab("Latitude")+ggtitle("Without Daisies")
```
]
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
plot
```
---
.small[
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
plot2 <- ggplot(data.frame(I,Barr,Warr,Uarr,Tarr))+geom_line(aes(I,Barr),color="black")+ geom_line(aes(I,Warr),color="white")+ geom_line(aes(I,Uarr),color="brown")+geom_line(aes(I,Tarr/25),color="pink")+ylab("Temperature & Albedo")+xlab("Time")+ggtitle("Without Daisies")
```
]
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
plot2
```
---
.small[
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
plot_ly(z=~TEMP)%>% add_surface() %>% layout(title="Without Daisies",
  scene = list(
    xaxis = list(title = "Time"),
    yaxis = list(title = "Latitude"),
    zaxis = list(title = "T")  )) 
```
]
---
## EBM without Daisies
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
a <- w*aW+b*aB+u*alb(T,0.62,0.25)
T <- gauss(Zones,0,50,31.6)-6
TEMP <- matrix(NA, nrow=90, ncol=500)
for(i in c(1:500)) {
S <- Sun6(i)  
Rin <- Incident(S,SunWt)
Tcos <- cosZones*T
Tm <- sum(Tcos)/sum(cosZones)
T <- (Rin*(1-a)+K*Tm-A) / (B+K)
TEMP[,i] <- T
Tw <- T+c*(a-aW)
Tb <- T+c*(a-aB)
Fw <- 1-k*(T0-Tw)^2
Fb <- 1-k*(T0-Tb)^2
for(j in c(1:length(Zones))){ if(Fw[j]<0){Fw[j]=0}
  if(Fb[j]<0){Fb[j]=0}  }
w <- w+w*(u*Fw-D)
b <- b+b*(u*Fb-D)
for(j in c(1:length(Zones))){if(w[j]<0.001){w[j]=0.001}
if(b[j]<0.001){b[j]=0.001}  }
u <- 1-w-b
a <- w*aW+b*aB+u*alb(T,0.62,0.25)
Barr[i] <- b[45]
Warr[i] <- w[45]
Uarr[i] <- u[45]
I[i] <- i
Tarr[i] <- T[45]
} 
```
---
.small[
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
plot <- ggplot(data.frame(Zones,w,b,u,T))+geom_line(aes(Zones,w),color="white")+geom_line(aes(Zones,b),color="black")+geom_line(aes(Zones,u),color="brown")+geom_line(aes(Zones,T/5),color="green")+ylab("Temperature & Albedo")+xlab("Latitude")+ggtitle("With Daisies")
```
]
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
plot
```
---
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
plot2 <- ggplot(data.frame(I,Barr,Warr,Uarr,Tarr))+geom_line(aes(I,Barr),color="black")+ geom_line(aes(I,Warr),color="white")+ geom_line(aes(I,Uarr),color="brown")+geom_line(aes(I,Tarr/25),color="pink")+ylab("Temperature & Albedo")+xlab("Time")+ggtitle("With Daisies")
```
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
plot2
```
---
.small[
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
plot_ly(z=~TEMP)%>% add_surface() %>% layout(title="With Daisies",
  scene = list(
    xaxis = list(title = "Time"),
    yaxis = list(title = "Latitude"),
    zaxis = list(title = "T")  )) 
```
]



